Applying Multi-Core Model Checking to Hardware-Software Partitioning in Embedded Systems

### 2015-09-01 Docs/Paper/LaTex/template.tex
2.4 Solving Optimization Problems with Vz

An SMT solver is a tool that verifies formula satisfiability based on a set of background theories ({\it e.g.}, linear arithmetic, bitvectors, and arrays). \textcolor{red}{Z3 is built on the a Davis–Putnam–Logemann–Loveland algorithm together with auxiliary algorithms for solving quantifiers formulas~\cite{Bjorner2014}}. vZ is one of the tools implemented in the SMT solver Z3; its base function is to optimize objective functions, which formulate optimized criteria, within the logical context of constraints.  vZ includes an incremental version of  \textcolor{red}{Maximum Resiliency (MaxRes) ~\cite{Federica2008} to Maximum Satisfiability (MaxSAT)~\cite{NarodytskaN}} and a Simplex to solve numbers without defined patterns. In Vz, MaxSAT is responsible for the restrictions, while OptSMT optimizes linear arithmetic objectives~\cite{Bjorner2015}. Vz provides functions that extend Z3 for solving optimization problems: \textit{maximize}, \textit{minimize}, and \textit{assert-soft}.

\begin{itemize}
\item{\textbf{maximize(T)}
this function informs to the solver that a given variable $T$ should be maximized, which includes real, integer, or bit-vector variables.}
\item{\textbf{minimize(T)}
this function informs the solver that a given variable $T$ should be minimized, the accepted types are the same as maximize function.}
\item{\textbf{Assert-Soft F : weight n}
the function Assert-Soft adds a restriction to $F$, which can also add a weight $n$; the default value is $1$.}
\end{itemize}

@@

\begin{lstlisting}[caption=SMT formula using vZ, label=vZ]
1. (declare-const K Int) 
2. (declare-const W Int)
3. (assert (< K 2)) 
4. (assert (< (- W K) 1))
5. (maximize (+ K W)) 
6. (check-sat)
\end{lstlisting}

The vZ architecture is shown in Fig.~\ref{vZ-Architecture}. Initially, the SMT formula with objectives is converted to $0-1$ constraints, which leads to a Pseudo-Boolean Optimization (PBO)\textcolor{red}{~\cite{Vasco2005}~\cite{Barth1995}}. If there are many objective functions, vZ invokes OptSAT for arithmetic or MaxSAT for soft constraints. \textcolor{red}{for box constraints using real values, Vz combines linear arithmetic objectives and uses only one instance of OptSMT. When we use "soft constrains'' in the mode "lexicographic'' vZ invokes MaxSAT using multiple calls for your engine.} 
%
\begin{figure}[ht]
	\centering
  \includegraphics[width=0.49\textwidth, height=95px]{Image/vzArch.png} 
	\caption{vZ Architecture~\cite{Bjorner2015}}
	\label{vZ-Architecture}
\end{figure}

Z3 is available for platforms in C, C$++$, Java, .NET and Python; it is possible to download Z3 with vZ from~\cite{Z3API}. In this work, the python API is used to formulate the HW-SW partitioning problem in vZ tool.

>>>>

SMT decides the satisfiability of first-order formulas using a combination of different background theories and thus generalizes propositional satisfiability by supporting uninterpreted functions, linear and nonlinear arithmetic, bitvectors, tuples, arrays, and other decidable first-order theories. In this study, the SMT solver Z3 is used to check for the satisfiability of formulas generated from the HW-SW partitioning problem~\cite{Bjorner2014}. In particular, we exploit the use of vZ, which is implemented on top of the SMT solver Z3, in order to solve optimization problems; Vz base function is to optimize objective functions, which formulate optimized criteria, within the logical context of constraints.~vZ also includes an incremental version of the Maximum Resiliency (MaxRes) \cite{Federica2008} in order to achieve Maximum Satisfiability (MaxSAT) \cite{NarodytskaN} and a Simplex to solve numbers without defined patterns. In Vz, MaxSAT is responsible for the restrictions, while OptSMT optimizes linear arithmetic objectives~\cite{Bjorner2015}. In summary, Vz provides three main functions that extend Z3 for solving optimization problems, which are: \textit{maximize}, \textit{minimize}, and \textit{assert-soft}.

\begin{itemize}
\item{\textbf{maximize(T)}
this function informs to the solver that a given variable $T$ should be maximized, which includes real, integer, or bit-vector variables.}
\item{\textbf{minimize(T)}
this function informs the solver that a given variable $T$ should be minimized, the accepted types are the same as maximize function.}
\item{\textbf{Assert-Soft F : weight n}
the function \textit{assert-soft} adds a restriction to $F$, which can also add a weight $n$; the default value is $1$.}
\end{itemize}

@@

\begin{lstlisting}[caption=Example of SMT formula using vZ, label=vZ]
1. (declare-const K Int) 
2. (declare-const W Int)
3. (assert (< K 2)) 
4. (assert (< (- W K) 1))
5. (maximize (+ K W)) 
6. (check-sat)
\end{lstlisting}

The vZ architecture is shown in Fig.~\ref{vZ-Architecture}. Initially, the SMT formula with objectives is converted to $0-1$ constraints, which leads to a Pseudo-Boolean Optimization (PBO)~\cite{Barth1995,Vasco2005}. If there are many objective functions, vZ invokes OptSAT for arithmetic or MaxSAT for soft constraints. \textcolor{red}{Qual seria o ``box constraints''?} For box constraints using real values, Vz combines linear arithmetic objectives and uses only one instance of OptSMT. When ``soft constrains'' is used in the mode `` lexicographic'', vZ invokes MaxSAT using multiple calls for its engine. \textcolor{red}{Qual ``engine''?} 
%
\begin{figure}[ht]
	\centering
  \includegraphics[width=0.49\textwidth, height=95px]{Image/vzArch.png} 
	\caption{vZ Architecture~\cite{Bjorner2015}}
	\label{vZ-Architecture}
\end{figure}

Z3 is available for platforms in C, C$++$, Java, .NET, and Python; it is possible to download Z3 with vZ from its github repository~\cite{Z3API}. In this work, the python API is used to formulate HW-SW partitioning problems using the vZ tool.
